### 2014-12-03 manuscript/manuscript.Rnw
First commit. One figure added and part of the abstract.

\documentclass[a4paper]{article}

%--- Load packages. 
\usepackage{amssymb,amsmath}
\usepackage{framed}

% avoid ugly indentation of paragraphs
\usepackage{parskip}
\usepackage{graphicx}

% open sans font
\usepackage[default,osfigures,scale=0.95]{opensans}

% subfolder with the figures (within manuscript/)
\graphicspath{ {./figures/} }

% for author affiliations
\usepackage{authblk}

% for degree symbol
\usepackage{gensymb}

% for line numbers
\usepackage{lineno}

% margin size
\usepackage{geometry}
\geometry{verbose,tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}

% for bibtex
\bibliographystyle{plain}

\begin{document}

% Turn line numbers on here.
%\linenumbers

\title{Catchy article title}

\author[1]{Remko Duursma}

\affil[1]{Hawkesbury Institute for the Environment}

\renewcommand\Authands{ and }
\maketitle

%--------------------------------------------------------------------------------------------%

<<preload, echo=FALSE>>=
library(knitr)
opts_chunk$set(echo=FALSE)
@

%--------------------------------------------------------------------------------------------%

\section*{Summary}

1. The \texttt{plantecophys} R package includes many functions for analysis and modelling of plant leaf gas exchange data.

2. This paper describes the new package, and gives examples on its use in plant ecophysiological research. The package makes technically difficult calculations easily accessible for many users.

3. We briefly describe functions for modelling C3 leaf gas exchange with the coupled photosynthesis-stomatal conductance model, fitting of A-Ci curves, modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation.

\section*{Introduction}

Time to cite some papers, some folks said so \cite{schenk_rooting_2002}, and some others as well \cite{mokany_critical_2006}.

%--------------------------------------------------------------------------------------------%

\clearpage
\section{Figures}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.99\textwidth]{figure1.pdf}
    \caption{Standard output from the \texttt{fitaci} function. $A_net$ is the net photosynthetic rate, $C_i$ the intercellular CO$_2$ concentration. Symbols are measurements, the line the fitted Farquhar model of photosynthesis.}
    \label{fig:figure1}
\end{figure}

\clearpage
\bibliography{references}

\end{document}