### 2015-07-09 manuscript.Rmd
Final polishing; as submitted to PLOS ONE 2015-07-09

---
title: "plantecophys - an R package for analysing and modelling leaf gas exchange data"
author: ''
date: ''
output:
  word_document:
    reference_docx: manuscriptstyle.docx
  pdf_document: default
  html_document:
    number_sections: yes
csl: plos-one.csl
bibliography: references.bib
---

\
Remko A. Duursma^1^

\
^1^ Hawkesbury Institute for the Environment, University of Western Sydney, Locked Bag 1797, Penrith, NSW, Australia

\
*Corresponding author*: 
Remko Duursma
E: remkoduursma@gmail.com
T: +61(0)422096908

*Short title*: The plantecophys R package

<!---
For Submission to PLOS one
e.g. http://journals.plos.org/plosone/article?id=10.1371/journal.pone.0061217
http://journals.plos.org/ploscollections/article?id=10.1371/journal.pcbi.1003929
http://journals.plos.org/ploscollections/article?id=10.1371/journal.pcbi.1003270
and see: http://journals.plos.org/plosone/s/submission-guidelines
(Methods, software, databases, and tools)
--->

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Two useful packages for printing numbers, accessing content of objects.
library(broom)
library(reporttools)

# Tables in markdown
library(ascii)

# mine
library(plantecophys)

# use git
library(git2r)

# figures
source("figures.R")

# output
if(!dir.exists("output"))dir.create("output")

# Do calculations here, create objects that we can use throughout the paper.

# Or, place them all in a separate script:
# source("manuscript_calcs.R")

# Shorthand function to print numbers with two decimal points (and retain trailing zeroes)
f <- function(x, digits=2)sprintf(x, fmt=paste0("%.",digits,"f"))

# Nice function to format P values.
pval <- function(...)formatPval(..., includeEquality=TRUE)

# Set Flag to include figures in text or not.
includeFigs <- FALSE

# This document can be compiled with the 'Knit Word' button in Rstudio, or:
# library(rmarkdown)
# render("manuscript.Rmd", "word_document", "manuscript.docx")

# Find current SHA of plantecophys
rep_pkg <- "c:/repos/plantecophys"
if(file.exists(rep_pkg)){
  pkgrep <- repository(rep_pkg)
  r <- reflog(pkgrep, verbose=FALSE)
  cur_sha <- substr(r[[1]]@sha, 1, 7)
} else {
  cur_sha <- "???"
}

# Make PDFs
to.pdf(figure1(), "output/figure1.pdf", width=8, height=4, verbose=FALSE)
to.pdf(figure2(), "output/figure2.pdf", width=4, height=4, verbose=FALSE)
to.pdf(figure3(), "output/figure3.pdf", width=8, height=4, verbose=FALSE)
to.pdf(figure4(), "output/figure4.pdf", width=8, height=4, verbose=FALSE)
```

# Abstract {.unnumbered}

Here I present the R package 'plantecophys', a toolkit to analyse and model leaf gas exchange data. Measurements of leaf photosynthesis and transpiration are routinely collected with portable gas exchange instruments, and analyzed with a few key models. These models include the Farquhar-von Caemmerer-Berry (FvCB) model of leaf photosynthesis, the Ball-Berry models of stomatal conductance, and the coupled leaf gas exchange model which combines the supply and demand functions for CO~2~ in the leaf. The 'plantecophys' R package includes functions  for fitting these models to measurements, as well as simulating from the fitted models to aid in interpreting experimental data. Here I describe the functionality and implementation of the new package, and give some examples of its use.  I briefly describe functions for fitting the FvCB model of photosynthesis to measurements of photosynthesis-CO~2~ response curves ('A-C~i~ curves'), fitting Ball-Berry type models, modelling C3 photosynthesis with the coupled photosynthesis-stomatal conductance model, modelling C4 photosynthesis, numerical solution of optimal stomatal behaviour, and energy balance calculations using the Penman-Monteith equation. This open-source package makes technically challenging calculations easily accessible for many users and is freely available on CRAN.

\
Keywords: photosynthesis, stomatal conductance, portable gas exchange instrument, process-based model, software

# Introduction

Since the advent of portable gas exchange instruments [@mcdermitt1989; @leuning1989], a wealth of data on leaf gas exchange of CO~2~ and H~2~O has been collected  [@long2003]. These data play a central role in physiological plant ecology [@pearcey1989], to better understand and quantify inter-specific differences in photosynthesis and transpiration, and to quantify and model the rapid response to changes in environmental drivers such as light, humidity and temperature. Not only do leaf gas exchange data allow detailed studies of the underlying plant physiology, they are also used to parameterize an important component of process-based models of vegetation function used to predict global water and carbon cycling [e.g. @bonan2014; @dekauwe2015].

\
The photosynthesis model of Farquhar, von Caemmerer and Berry [@farquhar1980] (the 'FvCB model') is widely used in interpreting and modelling leaf gas exchange, by providing comparable metrics of the photosynthetic capacity, and predicting the response of photosynthesis to changes in the CO~2~ concentration inside the leaf air space (C~i~). This widely cited model is embedded in many process-based models of vegetation function [@sellers1997; @bonan2014]. The key prediction of the model is the response of photosynthesis to [CO~2~] inside the leaf (either chloroplastic [CO~2~], C~c~, or intercellular [CO~2~], C~i~). It can also account for changes in leaf temperature if the various temperature sensitivities are parameterized [@caemmerer1981; @medlyn2002]. To employ the model, it is generally fit to observations of net photosynthesis along a range of [CO~2~] concentrations, yielding well-known measures of photosynthetic capacity (V~cmax~ and J~max~) [@wullschleger1993].

\
We do not repeat a detailed description of the FvCB model here, as it has been described many times [e.g. @medlyn2002]. But generally it is of the form,

\
(1)
$$A_n = min(A_c, A_j) - R_d$$

\
where A~n~ is the net rate of CO~2~ assimulation, A~c~ is the gross photosynthesis rate when Rubisco activity is limiting, A~j~ when RuBP-regeneration is limiting, and R~d~ the rate of dark respiration (see Fig. 1a). A~c~ and A~j~ are non-linear functions of the intercellular CO~2~ concentration (C~i~), both of the form $k_1(C_i{\Gamma}^{∗})/(k_2 + C_i)$, where ${\Gamma}^{∗}$ is the CO~2~ compensation point without R~d~, and k~1~ and k~2~ are different parameter combinations for A~c~ and A~j~. The details of these functions and the temperature dependence of the various parameters are described elsewhere [@medlyn2002]. We have here used C~i~ to denote [CO~2~] concentration in the leaf. Alternatively we may account for the diffusional limitation from intercellular airspaces to the chloroplast [@ethier2004], by calculating C~c~ as,

\
(2)
$$C_c = C_i - A_n/g_m$$

where g~m~ is the mesophyll conductance.

\
**Figure 1**. (a) Leaf photosynthesis - CO~2~ response curve as modelled with the FvCB model. (b) The intersection of the supply and demand curves of photosynthesis. The `Photosyn` function solves for C~i~ if g~s~, V~cmax~, J~max~ and R~d~ (and other parameters to the FvCB model) are known.

\
Fitting the model to data requires some finesse because the model is a minimum function of two non-linear equations that is sometimes difficult to fit. Moreover, sample sizes collected are often small due to time constraints in the field. A widely used published method requires the user to specify the transition of V~cmax~ to J~max~ limitation [@sharkey2007], a process that is both arbitrary and prevents batch analysis. Another method [@leafweb] requires online submission of data and fits the model without much control or knowledge of the fitting process (following [@gu2010]). Undoubtedly many more implementations of the fitting process have been developed over the years, but few of these are made publicly available (but see available online tools [@dekauwefitfarquhar; @landfluxtools]). What is missing is an open-source tool that can be used for reproducible and transparent analysis of A-C~i~ curves.

\
Through Eq. 1, we have a dependency of photosynthesis on the availability of the substrate, C~i~. To estimate C~i~ itself, we need an estimate of the stomatal conductance to CO~2~. From Fick's law, we can relate A~n~ to g~s~ and C~i~ as,

\
(3)
$$ A_n = g_s(C_a - C_i) $$

\
where g~s~ is the conductance to CO~2~. We now have two equations for A~n~. The 'demand function' is given by Eq. 1, and the 'supply function' in Eq. 3. At steady state these two equation should be equal, which can be graphically shown as in Fig. 1b (cf. [@farquhar1982]).

\
Because g~s~ itself responds to environmental drivers, another expression is needed to end up with a fully coupled model of leaf gas exchange. The most widely-used, though empirical, g~s~ model is the Ball-Berry [@ballberry1987] class of models. This model posits an entirely empirical equation that describes the response of g~s~ to air humidity, CO~2~ and A~n~. This way, effects of leaf temperature and PPFD - both of which are known to affect g~s~ - are modelled through the dependency of A~n~ on these drivers. A general form of the Ball-Berry model is,

\
(4)
$$ g_s = g_0 + g_1 \frac{A_n}{C_a}f(D)$$

\
where D is the vapour pressure deficit (kPa), g~0~ and g~1~ are empirical parameters, and f(D) a function that describes the response to the vapour pressure deficit (D, [@leuning1995]) or relative humidity [@ballberry1987]. An alternative approach to modelling g~s~ is through the hypothesis that stomata act optimally in the sense that they maximize photosynthesis while minimizing water loss. This hypothesis was first developed by Cowan and Farquhar [@cowan1977] and has seen many applications. Medlyn et al. [@medlyn2011] showed that the optimality hypothesis, when coupled to the FvCB model, leads to an expression analogous to the Ball-Berry type models (Eq. 4), but with a different D response function (f(D) in Eq. 3) than the original Ball-Berry model.

\
Finally we can combine the biochemical demand function of photosynthesis (Eq. 1) with the supply function (Eq. 2) and an expression for the dependency of g~s~ on environmental drivers (Eq. 3). This 'coupled' leaf gas exchange model [@leuning1990; @collatz1991; @leuning1995] is implemented in many process-based ecosystem and global land surface models [@wang1998; @bonan2014; @duursma2012; @sellers1997]. This model allows prediction of A~n~, g~s~ and leaf transpiration rate in response to all major environmental drivers (except soil water limitation), and key leaf traits (g~1~, V~cmax~, J~max~, R~d~, and their temperature dependencies).

\
Despite the widespread use of the FvCB model and the coupled leaf gas exchange model, tools to analyze data and perform simulations are scattered and subject to little standardization. Fitting the FvCB model to CO~2~ response curves is a standard procedure but different methods can yield different parameter values, making comparisons difficult. The coupled leaf gas exchange model is not straightforward to implement, and I do not know of any standalone open-source implementations. I here describe the plantecophys package, implemented in the R language [@rfoundation]. The code is freely available (without restrictions), and managed with a version control system. The package is the result of our work on leaf and canopy modelling of photosynthesis and stomatal conductance [@medlyn2011; @barton2012; @peltoniemi2012; @medlyn2013; @duursma2013; @duursma2014; @lin2015], with many additions based on user requests.

# Design and implementation

## The main functions

The main tools included in the plantecophys package are to a) fit A-C~i~ curves, b) fit Ball-Berry type models, c) simulate from the coupled leaf gas exchange model and d) calculate the optimal stomatal conductance. A list of the key functions in the package are summarized in Table 1.

```{r results='asis', echo=FALSE, message=FALSE, warn=FALSE}
f <- c("`fitaci`","`fitBB`","`FARAO`","`Photosyn`","`PhotosynEB`","`AciC4`", "`RHtoVPD` etc.")
d <- c("Fit, summarize, plot and simulate photosynthesis-[CO~2~] response curves (A-Ci curves)",
"Fit Ball-Berry type models of stomatal conductance",
"Estimate optimal stomatal conductance with a numerical implementation of the Cowan-Farquhar hypothesis",
"Simulate C3 photosynthesis and transpiration with the coupled leaf gas exchange model. Also simulates the FvCB model when either C~i~ or g~s~ is given as input. For C4 photosynthesis, see `AciC4`",
"Estimate leaf temperature from energy balance, when a significant leaf boundary layer is present",
"Simulates the dependence of C4 photosynthesis on the intercellular CO~2~ concentration",
"Convert between commonly used units (relative humidity, vapour pressure deficit, dewpoint temperature)")
df <- data.frame(Function=f, Description=d)

suppressWarnings(print(ascii(df, caption="Table 1. Main functions in the plantecophys package.",
            include.rownames=FALSE),
      type='pandoc'))
```

## Language

The 'plantecophys' package is implemented in R, has no dependencies on other packages, and does not require compilation (i.e. it is written in native R only). As such it builds easily, and is highly portable. The source code is maintained with git version control, and is hosted in an online repository (http://www.bitbucket.org/remkoduursma/plantecophys). The repository includes an issue tracker, where users can suggest changes or report bugs.  This paper describes version `r packageVersion("plantecophys")` (git SHA `r cur_sha`).

## Results and Discussion

### Fitting A-C~i~ curves

The `fitaci` function fits the FvCB model, yielding estimates of V~cmax~, J~max~ and R~d~ and their standard errors. Instead of fitting the minimum function (Eq. 1), `fitaci` fits the hyperbolic minimum of A~c~ and A~j~, which avoids a discontinuity. The fit is achieved via standard non-linear least squares, and standard errors of the parameters are estimated with standard methods (`nls` function in base R). The `fitaci` function includes methods to estimate appropriate starting values from the data, and attempts the fits along a wide range of possible starting values. Optionally, R~d~ can be provided as a known value, otherwise it is estimated from the A-C~i~ curve. The user does not have to provide the transition point (see Fig. 1), as this is estimated by `fitaci` automatically. It is however an option to fix the transition point (via the `citransition` argument), which may be helpful to check whether the best fit was achieved. Finally, the user can provide an estimate of mesophyll conductance (g~m~) (following [@ethier2004]), in which case the fitted values of V~cmax~ and J~max~ can be interpreted as chloroplastic rates.

\
In the FvCB model, J~max~, V~cmax~ and leaf respiration (R~d~) (and other parameters like $\Gamma^{*}$, K~c~ and K~o~) all depend non-linearly on T~leaf~. The `Photosyn` function incorporates standard temperature sensitivities for all parameters of the FvCB model (following [@medlyn2002]). Optionally, measured K~m~ and $\Gamma^{*}$ can be provided instead.

\
The function takes a dataframe as input, which includes measurements of A~n~, C~i~ and optionally leaf temperature (T~leaf~) and photosynthetically active radiation (PAR). It then returns a special object which has a print and plot method. It is easily used like this:

```
# Fit FvCB model
f <- fitaci(mydata)

# Print a summary with coefficients and more
f

# Make standard plot
plot(f)
```

\
The output of the above example is shown in Fig. 2. Additionally, the batch utility `fitacis` can be used to fit many curves at once, for example one for each species or site in a dataset.

\
**Figure 2**. Standard output from the `fitaci` function. A~n~ is the net photosynthetic rate, C~i~ the intercellular CO~2~ concentration. Symbols are measurements, the black line the fitted FvCB model of photosynthesis. Colored lines indicate the two photosynthesis rates in the FvCB model.

\
A C4 model of leaf photosynthesis [@caemmerer2000] is also implemented (in `AciC4`), but at the moment it is only possible to fit the C3 model of leaf photosynthesis to A-C~i~ curves.

### Fitting stomatal conductance models

The straightforward `fitBB` function provides an interface to non-linear or linear regression to fit one of three stomatal conductance models [@ballberry1987; @leuning1995; @medlyn2011]. This yields estimates of g~1~ and (optionally ) g~0~, which are necessary inputs to the coupled leaf gas exchange model. Note that, unlike the description in this paper, the user must provide stomatal conductance to H~2~O (not CO~2~) as input to the fitting process, which is the standard output of portable gas exchange instruments.

### Coupled leaf gas exchange model

The intersection of the supply and demand curves of photosynthesis (Fig. 1b) gives the steady-state intercellular CO~2~ concentration. This is solved by the `Photosyn` function. This flexible interface can be used to either 1) estimate A~n~ when C~i~ is known (`Photosyn(Ci=...)`, equivalent to `Aci(...)`), 2) estimate A~n~ when g~s~ is known (`Photosyn(GS=...)`) (cf. Fig. 1b) or c) solve for C~i~ from the coupled leaf gas exchange model (Eqs. 1,3,4).

\
To demonstrate the use of the coupled gas exchange model, I visualize the temperature response of A~n~ when both T~leaf~ and D are varying. In field conditions, D is always strongly positively related to T~leaf~. The consequence is that when studying D or T~leaf~ responses in the field, both drivers have to be accounted for [@lin2012; @duursma2014]. Figure 3a shows simulated A-C~i~ curves and the solutions of the coupled leaf gas exchange models at a range of T~leaf~ and corresponding D (calculated following [@duursma2014]). Figure 3b shows the output from `Photosyn`, which can be created with the following example. Note that in this example the default values of many parameters (e.g. J~max~, g~1~) are used in the call to `Photosyn`, but all of these can be set by the user.

\
**Figure 3**. Response of A~n~ and C~i~ to combined changes in T~leaf~ and D. (a) Lines are A-C~i~ curves simulated at a range of values for T~leaf~. Symbols are the solutions of the coupled leaf gas exchange model, while also taking into account the correlation between D and T~leaf~ (based on an empirical relationship [@duursma2014]). Note that as T~leaf~ and D increase, C~i~ decreases. (b) The corresponding temperature optimum of A~n~. Symbols are the same as in panel (a) but plotted against T~leaf~.

```
# Set range of leaf temperature
tleafs <- seq(5, 40, by=5)

# Define D as a function of Tleaf
vpdfun <- function(tair)0.000605*tair^2.39

# Simulate.
run1 <- Photosyn(Tleaf = tleafs, VPD=vpdfun(tleafs))

# Plot (produces Figure 3b minus the formatting)
with(run1, plot(Tleaf, ALEAF))

```

\
The `Photosyn` function assumes that the boundary layer conductance (g~bl~) is high compared to g~s~, so that T~leaf~ is close to T~air~. As an alternative, the `PhotosynEB` function calculates T~leaf~ from the leaf energy balance. Transpiration is calculated with the Penman-Monteith equation, which accounts for boundary layer effects. The details of `PhotosynEB` are not described here (we refer to the documentation), because it is very similar to other implementations [@wang1998; @buckley2014].

### Numerical solution of optimal stomatal conductance

The `FARAO` function (FARquhar And Optimality) calculates optimal stomatal conductance based on the Cowan-Farquhar hypothesis. It finds the C~i~ for which the quantity $A - \lambda E$ is maximal, and then calculates corresponding A~n~ and g~s~. This implementation was used by Medlyn et al. [@medlyn2011] to compare a simplified model of optimal stomatal conductance to the full numerical solution. In Fig. 4a, I have calculated $A - \lambda E$ across a range of C~i~ values, and for different values of D. The `FARAO` function calculates the optima of these curves, which can for example be used to study the response of stomatal conductance to D (Fig. 4b).

\
**Figure 4**. Visualization of the optimal model of stomatal conductance. Provided we have an estimate of the 'cost of water' ($\lambda$, mol C mol H~2~O^-1^), stomata act to maximize photosynthesis minus transpiration. In (a), individual curves at a range of values for the vapour pressure deficit (D) plots $A-\lambda E$ versus the C~i~, demonstrating that an optimum exists. The `FARAO` function finds this optimum numerically and calculates A~n~ and g~s~. The corresponding response of g~s~ to D is shown in panel (b).

\
Optionally, the `FARAO` function accounts for the presence of a leaf boundary layer (when `energybalance=TRUE`). In that case it uses `PhotosynEB` (see description above) to calculate A~n~ and E, and solves for T~leaf~. A very similar method was employed by Buckley et al. [@buckley2014], who demonstrated that when a boundary layer is present, frequently an optimal g~s~ cannot be found.

# Conclusions

We need an open source set of tools to analyze leaf gas exchange data, as these data form a cornerstone of plant physiological ecology. At the moment there are no publicly available tools to fit A-C~i~ curves that can be used as part of a reproducable workflow. The plantecophys R package is implemented in widely used language for data analysis. The package includes a useful set of tools to perform standard, and more advanced, analyses of leaf gas exchange data. The open source framework combined with version control allows further development of the code.

# Availability and requirements

**Project name**: plantecophys

**Project Stable Release**: cran.r-project.org/package=plantecophys

**Project Home Page**: http://www.bitbucket.org/remkoduursma/plantecophys

**Project Issue Tracker**: http://www.bitbucket.org/remkoduursma/plantecophys/issues

**Operating System(s)**: Platform Independent

**Programming Language(s)**: R

**Other Requirements**: none

# Acknowledgments {.unnumbered}
David Ellsworth, Belinda Medlyn and Martin de Kauwe are acknowledged for discussions on fitting A-C~i~ curves. Special thanks to Belinda Medlyn whose code in the MAESTRA model provided the basis for an early version of the Photosyn function. Thanks to John Drake for testing.

# References {.unnumbered}